Diffusion Models for Tissues and Tumor Growth

### 2024-03-17 paper.tex
one other typo

Researchers have been studying cancer and the patterns behind tumor growth for thousands of years. Hippocrates
was accredited with possibly being the first to describe a tumor, which were mainly described by their color
and their response to touch. However, ancient physicians did not separate benign and malignant tumors, but they
acknowledged the tumor's ability to infect nearby tissue. The application of mathematical models to tumor growth
did not start to gain traction until much later, however. In the early 20th century, one of the most famous uses of
the diffusion model was first developed: Hill released a paper discussing how the movement of lactic acid
and oyxgen throughout the body could be shaped into a diffusion problem (Hill, 1928). This paper eventually led to
Greenspan's paper which modeled the growth of tumors using diffusion, applying many of the mathematical tools employed
by Hill some years before (Greenspan, 1972). This paper will present an overview of the diffusion model in mathematical
biology and will discuss the mathematics and biology behind two of the most influential diffusion papers.

@@

Let $x$ be the distance from the constant source of oxygen to any point within the tissue. Given the assumptions above,
the rate of diffusion across any unit area will be $ J = -k \* \frac{dy}{dx}$,\, where $J$ is the rate of diffusion, $k$ is the diffusion rate and the (-)
represents that the diffusion is occuring in the opposite direction from the source. This is more commonly known as Fick's
first law of diffusion. To find the measure of this rate of accumulation, Fick's second law can be applied:
\begin{equation}
    \frac{dJ}{dx} = k \* \frac{d^2y}{dx^2}
\end{equation}

>>>>

Researchers have been studying cancer and the patterns behind tumor growth for thousands of years. Hippocrates
was accredited with possibly being the first to describe a tumor, which were mainly described by their color
and their response to touch. However, ancient physicians did not separate benign and malignant tumors, but they
acknowledged the tumor's ability to infect nearby tissue. The application of mathematical models to tumor growth
did not start to gain traction until much later, however. In the early 20th century, one of the most famous uses of
the diffusion model was first developed: Hill released a paper discussing how the movement of lactic acid
and oxygen throughout the body could be shaped into a diffusion problem (Hill, 1928). This paper eventually led to
Greenspan's paper which modeled the growth of tumors using diffusion, applying many of the mathematical tools employed
by Hill some years before (Greenspan, 1972). This paper will present an overview of the diffusion model in mathematical
biology and will discuss the mathematics and biology behind two of the most influential diffusion papers.

@@

Let $x$ be the distance from the constant source of oxygen to any point within the tissue. Given the assumptions above,
the rate of diffusion across any unit area will be $ J = -k \* \frac{dy}{dx}$,\, where $J$ is the rate of diffusion, $k$ is the diffusion rate and the (-)
represents that the diffusion is occuring in the opposite direction from the source. This is more commonly known as Fick's
first law of diffusion. To find the measure of this rate of accumulation, Fick's second law can be applied:
\begin{equation}
    \frac{dJ}{dx} = k \* (\frac{d^2y}{dx^2})
\end{equation}
